---
title: "Final Report"
subtitle: "BYUI - ECEN 490R"
author: "Kyle Tolliver"
date: "July 22, 2020"
---

<!-- You have been working this semester on a significant hardware or software project related to your major. You have submitted bi-weekly progress reports to document that you have adhered to your schedule and worked at least 100 hours on your project. -->

<!-- This week you will submit your Final Project Report. It must reflect the knowledge and skills expected from a senior-level student in your major. Those include writing skills as well as technical skills. In addition to being graded on your technical project work, you will be graded on your organization, your spelling, your grammar, and how well you followed the instructions for preparing your report. -->

<!-- Your Final Project Report must be at least six or eight single-spaced pages long, not including references. (It will take at least that much to adequately document your 100-hour project.) It must include the following sections and information. -->

```{r load_libraries, include=FALSE}
package::libraries()

wdset("D:/COVID-project/")

time <- read_csv("./time_logs/final.csv")
```

```{r wrangling}
dt <- time %>% 
  mutate(`Time (decimal)` = as.numeric(round(`Time (decimal)`, digits = 2))) %>% 
  rename(Time = `Time (decimal)`) %>% 
  select(Description, Time)
```

## Introduction

<!--
What problem or need does your project address? What does your project do? How does it solve the problem or fulfill the need? What has been done in the past by others to solve the problem or fulfill the need? Why is your project a better solution than what already existed?
-->

I am a senior majoring in Software Engineering, I plan to graduate in December of 2020. Additionally I am getting an Applied Associates in Data Science. So this project is a data science type project and includes a software design and development aspect. Some of the Data Science courses I have taken include the following:

* [Introduction to Databases](https://www.byui.edu/catalog#/courses/view/a7c647e4-9dbd-43ab-9f13-882ab279c65c)
* [Survey Object-Oriented Programming/Data Structures](https://www.byui.edu/catalog#/courses/4yzgxnx3ob?bc=true&bcCurrent=Survey%20Object-Oriented%20Programming%2FData%20Structures&bcGroup=Computer%20Science&bcItemType=Courses)
* [Data Intuition and Insight](https://www.byui.edu/catalog#/courses/view/5c868402f98eff24000cd2c4) 
* [Data Wrangling, and Visualization](https://www.byui.edu/catalog#/courses/view/5ce60ff8ecb90c2400a235c1) 
* [Database Design and Development](https://www.byui.edu/catalog#/courses/view/5c8baaf92679352400c8ad86)
* [Discrete Mathematics 1](https://www.byui.edu/catalog#/courses/view/5e1628b2f98ee825009625cc)
* [Calculus 1](https://www.byui.edu/catalog#/courses/4ybHLhehob?bc=true&bcCurrent=Calculus%20I&bcGroup=Mathematics&bcItemType=Courses)
* [Linear Algebra](https://www.byui.edu/catalog#/courses/Nyx8-l-3ib?bc=true&bcCurrent=Linear%20Algebra&bcGroup=Mathematics&bcItemType=Courses) 
* [Introduction to Web Design and Development](https://www.byui.edu/catalog#/courses/4ke4pgWniW?bc=true&bcCurrent=Introduction%20to%20Web%20Design%20and%20Development&bcGroup=Web%20Design%20and%20Development&bcItemType=Courses)

This project helped me strengthen the knowledge I gained from these classes, while learning new tools, techniques and technologies of Data Science and Web Development. 

I addressed the problem of unclear data on COVID-19. Many people see the data and can't make sense of it and what the values mean. I also figured how much of the population has been tested. I compared the techniques used to fight COVID-19, methods may include comparing the countries' data due to each using different methods to control the spread. My project helped people better understand COVID-19 precautions, influence and data. Each plot or data table has an impact on addressing the problem of unclear COVID-19 data. There are COVID-19 dashboards already created by multiple different companies. Two companies including Our World in Data and JHU CSSE have great dashboards already created. See [Dashboards](https://kctolli.github.io/COVID-19/final-report.html#References) under the references section, to get links to those dashboards.
  
My project wasn't just used for my personal learning, but also to provide useful data visualizations to help people understand COVID-19 and the data. My data presentation is a great solution due to the fact of using multiple different datasets to address the problem of COVID-19 data. Using different datasets allows people to compare and contrast the datasets. It also allows for new understanding and an expansion of data and visualizations.      

### Summary

* Developed a data website using R, Rmds and Yaml
* Explored different datasets to determine viable information
* Wrote my own COVID-19 R package to wrangle data
* Visualized the data to make the data easily understood

## Methods and Procedures

<!-- What research did you have to do? What are the project specifications? What tasks were required to implement your project? How long did they take? How close were your time estimates? What resources were needed? What was your test plan to verify that it worked correctly? -->

I found datasets that were useful for this project. I found three useful dataset hosts; they are NYTimes, Our World in Data and JHU CSSEGIS. See [Dataset References](https://kctolli.github.io/COVID-19/final-report.html#References) under the references section, to learn about the datasets used. The data wrangling and visualization is done in R. The data analytics is presented in a github hosted website. I used Rmd with R-chunks, and YML file to organize my data. I have many plots to test the same dataset and values.  

I used many R packages which include knitr, tidyverse, leaflet, sf (simple figures), dygraphs, USAboundaries, devtools, and more are used to implement the project. I also created my own R package to have functions used remotely. The package can be found on this projects [github repo](https://github.com/kctolli/COVID-19/tree/master/mypackage/package).  

Below is my completed work schedule and hour breakdown. I feel like I have was super close to my time estimates, but I worked more than the required time on this project which I expected. I was able to complete the each of the tasks in the specific interval.   

### Hour Breakdown

```{r time}
kable(dt)

x <- as.numeric(round(sum(dt$Time), digits = 2))

print(paste("Total hours equal:", x))
```

### Schedule

Below I have divided the tasks in three-week interval which a line progress reports.    
&emsp; I used [clockify](https://clockify.me/) to track hours. 

#### Whole Semester

Throughout the semester I spent time gaining knowledge on the follow as it relates to my project:  

* R packages
* Making Inferences with Data
* Analysis - Poor, [Good, Better, Best](https://www.churchofjesuschrist.org/study/general-conference/2007/10/good-better-best?lang=eng)
* Data Driven Learning
* Data Science Technologies 
* Resumes, Portfolios and Linkedin
* Engineering and Data Science Ethics 

#### Progress Report 1 (Due 5/9)

* Start Project
  + Proposal
* Do some data research
* Found data sets to work with
  + Such as [JHUs github](https://github.com/CSSEGISandData/COVID-19)
* Created github organization
* Hosted the website in main repo

#### Progress Report 2 (Due 5/30)

* Loaded in the data sets
  + Looked through the data set
* Started wrangling
* Made basic visualizations to see the data
* Developed YML file 

#### Progress Report 3 (Due 6/20)

* Wrangled and Visualized data
  + Non-interactive plots
  + Interactive plots
* Learned more about the 
  + [ggplot](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.0)
  + [sf plots](https://r-spatial.github.io/sf/articles/sf1.html)
  + [leaflet](https://leafletjs.com/)
  + [dygraphs](https://rstudio.github.io/dygraphs/)
  + other plot tools

#### Progress Report 4 (Due 7/11)

* Last website steps 
  + Finalize visualizations 
  + Finalize YML
  + Finalize tibbles
* Start report

#### Final Delivery (Due 7/22)

* Finish data presentation website
* Complete report

## Results

<!-- Include, as appropriate: schematics and wiring diagrams, assembly diagrams or photos, computer code, and results of the test plan (often including a screen shot or two, and a URL for a video). -->

I have created multiple different data visualizations to help address the problem of people unable to understand COVID-19 data. The website that presents the data visualizations is hosted on github: https://kctolli.github.io/COVID-19/index.html. This website shows visualizations for Johns Hopkins University and New York Times data.  

`maps <- leaflet(data = us) %>% addTiles()` 

`maps %>% `    
`  addPolygons(color = ~pal(cases), fillOpacity = 1, stroke = FALSE, group = "cases") %>%`    
`  addPolygons(color = ~pal(deaths),fillOpacity = 1, stroke = FALSE, group = "deaths") %>%`    
`  addPolygons(color = ~pal(per_cases), fillOpacity = 1, stroke = FALSE, group = "case %") %>%`    
`  addPolygons(color = ~pal(per_deaths), fillOpacity = 1, stroke = FALSE, group = "death %") %>%`    
`  addLayersControl(baseGroups = c("cases", "case %", "deaths", "death %"))`    

The visualization, I would like to highlight is found on the [JHU Leaflet page](https://kctolli.github.io/COVID-19/jhu_leaflet.html) and the code is displayed above. If you aren't familiar with leaflet, it is a R package that is used to make interactive maps. I have four overlay groups which are cases, case percent, deaths, death percent. There are colors for the overall data of COVID-19 for the group you select. The colors are green and red, with green being low value and red being high value, which is set with the pal variable `pal <- colorNumeric(c("green", "red"), NULL)`, and is implemented by the `~pal(value)`. The data of prop_states is found on this projects [github repo](https://github.com/kctolli/COVID-19/tree/master/read-data/JHU).

```{r map, fig.align='center', out.width='100%'}
map <- './ny_files/figure-html/mappinga-1.png'
include_graphics(map)
```

`ggplot(data = prop_state) +`    
`  geom_sf(aes(geometry = geometry), fill = NA, color = "Black") + guides(color = FALSE) + `    
`  geom_text(aes(label = (as.numeric(case_pop)*100), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +`    
`  theme_bw() +`    
`  theme(axis.title.x=element_blank(), axis.title.y=element_blank())` 

Above is another visualization, I'm going to highlight is found on the [New York Times page](https://kctolli.github.io/COVID-19/ny.html). The data of prop_states is found on this projects [github repo](https://github.com/kctolli/COVID-19/tree/master/read-data/nytimes). The code is also above used to create the map. The code is using the R packages sf (simple figures), USAboundaries and ggplot2 (part of the tidyverse package) to create the map.  

## Conclusions

<!-- How well does your project satisfy its intended purpose? How can it be applied or used by others? How can it be improved in the future? -->

My expected outcomes were to use data visualizations and documentation that is shown on a [github website](https://github.com/kctolli/COVID-19). I used have leaflets, dygraphs, ggplots and more visualization tools, to explain the data. The results are being shown using Rmd (RMarkdown) files that are put through knitr to get html. Yaml files were used to create the headers and javascript for the data website. This website helps people understand the data of COVID-19 using data tools such as visualizations. As people are aware and understand the data they can take appropriate actions. 

A few things I learned in this project. First, working with continuously updating data is hard unless you have an automatic run script. In the future I would setup a automatic run script to allow to always have up to date data. Second, sometime the provider of data corrupts the data. This I find is really strange, why would someone corrupt their own data? This happened in the process of this project, Our World In Data corrupted their own data causing my visualizations to stop working. This was only fixable by removing the data page from the website. Another thing I learned and implemented in the end of the project was to create an R package to create global functions that can be used in every page of the data website. I the future I would create the R package at the beginning, then update it as the project progresses. This motatived me to create my own general use [package](https://github.com/kctolli/Useful-R) full of useful functions. This project was a great learning experience for me and point to reflect on my data science skills.       

<br />

### Feedback

I was able to share my project with some family member and friends.                
They gave me some insist and feedback for the project and what to do moving forward.

* 

<br /> <br /> <br />

## References

<!-- Use APA format. -->

### Dashboards

Coronavirus Pandemic (COVID-19) - Statistics and Research. (2020). Retrieved 2020,            
&emsp; &emsp; from https://ourworldindata.org/coronavirus

Mapping COVID-19. (2020). Retrieved 2020,                                 
&emsp; &emsp; from https://systems.jhu.edu/research/public-health/ncov/

### Dataset References

CSSEGISandData. (2020). CSSEGISandData/COVID-19. Retrieved 2020,              
&emsp; &emsp; from https://github.com/CSSEGISandData/COVID-19

Nytimes. (2020). Nytimes/covid-19-data. Retrieved 2020,           
&emsp; &emsp; from https://github.com/nytimes/covid-19-data

Owid. (2020). Owid/covid-19-data. Retrieved 2020,               
&emsp; &emsp; from https://github.com/owid/covid-19-data

### Data Driven Learning

Babies - Netflix. (2020). Retrieved  2020,                
&emsp; &emsp; from https://www.netflix.com/watch/80117834?trackId=13752289

The Fallen of World War II - Data-driven documentary about war & peace. (n.d.). Retrieved 2020,             
&emsp; &emsp; from http://www.fallen.io/ww2/

### Additional Resourses

Grolemund, G. & Wickham, H. (n.d.). R for Data Science. Retrieved 2020,     
&emsp; &emsp; from https://r4ds.had.co.nz/

Hathaway, J. (n.d.). Consulting Articles. Retrieved 2020,      
&emsp; &emsp; from https://byuistats.github.io/M488/readings.html

Hathaway, J. (n.d.). R Visualization. Retrieved 2020,       
&emsp; &emsp; from https://byuistats.github.io/M335/r_visualization.html

Henry, L. & Wickham, H. (n.d.). Tidy evaluation. Retrieved 2020,       
&emsp; &emsp; from https://tidyeval.tidyverse.org/

Jenny Bryan, T. (n.d.). Happy Git and GitHub for the useR. Retrieved 2020,       
&emsp; &emsp; from https://happygitwithr.com/

Sievert, C. (2019, December 19). Data visualization with R, plotly, and shiny. Retrieved 2020,       
&emsp; &emsp; from https://plotly-r.com/index.html

Wickham, H. (n.d.). Advanced R. Retrieved 2020,       
&emsp; &emsp; from https://adv-r.hadley.nz/

Wickham, H. & Bryan, J. (n.d.). R Packages. Retrieved 2020,       
&emsp; &emsp; from https://r-pkgs.org/

Wilke, C. (n.d.). Fundamentals of Data Visualization. Retrieved 2020,       
&emsp; &emsp; from https://serialmentor.com/dataviz/

