---
title: "JHU Plots"
---

```{r setup, include=FALSE}
package::libraries()

wdset("D:/COVID-19/")

states <- states()
county <- county()
ny_county <- county() %>% newyork()

global <- my_read("./read-data/JHU/JHU_global.csv") 
us_pop <- my_read("./read-data/JHU/JHU_us_pop.csv")
us <- my_read("./read-data/JHU/JHU_usa.csv") 

us_state <- my_read("./read-data/JHU/JHU_us_state.csv") 
china_state <- my_read("./read-data/JHU/JHU_china_state.csv")
china <- my_read("./read-data/JHU/JHU_china.csv")
world <- my_read("./read-data/JHU/JHU_world.csv")
earth <- my_read("./read-data/JHU/JHU_earth.csv")
```

```{r us}
wide_us_cases <- us_state %>% select(- deaths) %>% 
  pivot_wider(names_from = state, values_from = cases) %>% 
  tk_xts(., date_var = date)

wide_us_deaths <- us_state %>% select(- cases) %>% 
  pivot_wider(names_from = state, values_from = deaths) %>% 
  tk_xts(., date_var = date)
```

```{r china}
wide_china_cases <- china_state %>% 
  select(- deaths, - recovered, - active) %>% 
  pivot_wider(names_from = state, values_from = cases) %>% 
  tk_xts(., date_var = date)

wide_china_deaths <- china_state %>% 
  select(- cases, - recovered, - active) %>% 
  pivot_wider(names_from = state, values_from = deaths) %>% 
  tk_xts(., date_var = date)

wide_china_recovered <- china_state %>% 
  select(- deaths, - cases, - active) %>% 
  pivot_wider(names_from = state, values_from = recovered) %>% 
  tk_xts(., date_var = date)

wide_china_active <- china_state %>% 
  select(- cases, - recovered, - deaths) %>% 
  pivot_wider(names_from = state, values_from = active) %>% 
  tk_xts(., date_var = date)
```

```{r world}
wide_world_cases <- world %>% 
  select(- deaths, - recovered, - active) %>% 
  pivot_wider(names_from = country, values_from = cases) %>% 
  tk_xts(., date_var = date)

wide_world_deaths <- world %>% 
  select(- cases, - recovered, - active) %>% 
  pivot_wider(names_from = country, values_from = deaths) %>% 
  tk_xts(., date_var = date)

wide_world_recovered <- world %>% 
  select(- deaths, - cases, - active) %>% 
  pivot_wider(names_from = country, values_from = recovered) %>% 
  tk_xts(., date_var = date)

wide_world_active <- world %>% 
  select(- cases, - recovered, - deaths) %>% 
  pivot_wider(names_from = country, values_from = active) %>% 
  tk_xts(., date_var = date)
```

## USA

### ggplots

```{r usa}
ggplot(data = us, aes(x = date, y = cases)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of USA's cases of COVID-19 by date. 

```{r usb}
ggplot(data = us, aes(x = date, y = deaths)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of USA's deaths of COVID-19 by date.

### dygraphs

```{r us_dya}
dygraph(wide_us_cases) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "cases")
```

This is an interactive graph of USA's cases of COVID-19 by date.

```{r us_dyb}
dygraph(wide_us_deaths) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "deaths")
```

This is an interactive graph of USA's deaths of COVID-19 by date.

## China

### ggplots

```{r chinaa}
ggplot(data = china, aes(x = date, y = cases)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of China's cases of COVID-19 by date.

```{r chinab}
ggplot(data = china, aes(x = date, y = deaths)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of China's deaths of COVID-19 by date.

```{r chinad}
ggplot(data = china, aes(x = date, y = recovered)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of China's recovered patients of COVID-19 by date.

```{r chinac}
ggplot(data = china, aes(x = date, y = active)) +
  geom_line() + 
  theme_bw()
```

This is a basic graph of China's active cases of COVID-19 by date.

### dygraphs

```{r china_dya}
dygraph(wide_china_cases) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "cases")
```

This is an interactive graph of China's cases of COVID-19 by date.

```{r china_dyb}
dygraph(wide_china_deaths) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "deaths")
```

This is an interactive graph of China's deaths of COVID-19 by date.

```{r china_dyc}
dygraph(wide_china_recovered) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "recovered")
```

This is an interactive graph of China's recovered patients of COVID-19 by date.

```{r china_dyd}
dygraph(wide_china_active) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "active")
```

This is an interactive graph of China's active cases of COVID-19 by date.

## World

The countries plotted are Australia, China, Denmark, France, Netherlands, and United Kingdom. 

### ggplots

```{r worlda}
ggplot(data = world, aes(x = date, y = cases, fill = country)) +
  geom_col() + facet_wrap(~country) +
  theme_bw()
```

This is a basic graph of multiple countries' cases of COVID-19 by date.

```{r worldb}
ggplot(data = world, aes(x = date, y = deaths, fill = country)) +
  geom_col() + facet_wrap(~country) + 
  theme_bw()
```

This is a basic graph of multiple countries' deaths of COVID-19 by date.

```{r worldd}
ggplot(data = world, aes(x = date, y = recovered, fill = country)) +
  geom_col() + facet_wrap(~country) +
  theme_bw()
```

This is a basic graph of multiple countries' recovered patients of COVID-19 by date.

```{r worldc}
ggplot(data = world, aes(x = date, y = active, fill = country)) +
  geom_col() + facet_wrap(~country) + 
  theme_bw()
```

This is a basic graph of multiple countries' active cases of COVID-19 by date.

### dygraphs

```{r world_dya}
dygraph(wide_world_cases) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "cases")
```

This is an interactive graph of multiple countries' cases of COVID-19 by date.

```{r world_dyb}
dygraph(wide_world_deaths) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "deaths")
```

This is an interactive graph of multiple countries' deaths of COVID-19 by date.

```{r world_dyc}
dygraph(wide_world_recovered) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "recovered")
```

This is an interactive graph of multiple countries' recovered patients of COVID-19 by date.

```{r world_dyd}
dygraph(wide_world_active) %>% 
  dyRangeSelector(height = 20, strokeColor = "") %>% 
  dyAxis("y", label = "active")
```

This is an interactive graph of multiple countries' active cases of COVID-19 by date.

<br /><br />

```{r update}
update()
```