---
title: "New York Times"
---

```{r setup, include=FALSE}
package::libraries()

wdset("D:/COVID-project/")

states <- states()
county <- county()
ny_county <- county() %>% newyork()

us <- my_read("./read-data/nytimes/nytimes_us.csv") %>% short_name()

state <- my_read("./read-data/nytimes/nytimes_state_pop.csv") %>% 
  mutate(`pop/mil` = pop/1000000)

state_prop <- my_read("./read-data/nytimes/nytimes_state_proportions.csv")

county_prop <- my_read("./read-data/nytimes/nytimes_county_proportions.csv")
```

```{r Wrangling}
prop_state <- left_join(state_prop, states, by = c("state", "state_abbr")) %>% 
  filter_yesterday() %>% select(- date) %>% covid_prop()

ny_county_prop <- county_prop %>% filter_yesterday() %>% newyork()

prop_county_ny <- left_join(ny_county_prop, ny_county, by = c("state", "state_abbr", "county")) %>% 
  select(- date, - state_abbr) %>% covid_prop()
```

## Find Distribution 

```{r find_top_complete}
ggplot(data = state) + 
  geom_col(aes(x = state, y = cases/pop)) +  
  coord_flip() + 
  theme_bw()

ggplot(data = state) + 
  geom_col(aes(x = state, y = deaths/pop)) + 
  coord_flip() + 
  theme_bw()
```

These plots visualize the top five states with covid cases and deaths per population.

## Maps

```{r mappingc}
ggplot(data = prop_state) +
  geom_sf(aes(geometry = geometry), fill = NA, color = "Black") + guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(cases)), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the exact value of cases from COVID-19.

```{r mappinga}
ggplot(data = prop_state) +
  geom_sf(aes(geometry = geometry), fill = NA, color = "Black") + guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(case_pop)*100), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the population that has tested positive from COVID-19.

```{r mappingd}
ggplot(data = prop_state) +
  geom_sf(aes(geometry = geometry), fill = NA, color = "Black") + guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(deaths)), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the exact value of deaths from COVID-19.

```{r mappingb}
ggplot(data = prop_state) +
  geom_sf(aes(geometry = geometry), fill = NA, color = "Black") + guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(death_pop)*100), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the population that has died from COVID-19.

### New York 

```{r newyorka}
ggplot(data = prop_county_ny) +
  geom_sf(aes(geometry = geometry), fill = NA,  color = "Black") +
  guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(case_pop)*100), geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 2) +
  theme_bw() + 
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the population that has tested positive for COVID-19 in the counties of New York.

```{r newyorkb}
ggplot(data = prop_county_ny) +
  geom_sf(aes(geometry = geometry), fill = NA,  color = "Black") +
  guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(death_pop)*100), geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 2) +
  theme_bw() + 
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the population that has died from COVID-19 in the counties of New York.

## Percentage of cases who died

```{r mappingz}
ggplot(data = prop_state) +
  geom_sf(aes(geometry = geometry), fill = NA,  color = "Black") +
  guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(death_case)*100), geometry = geometry), stat = "sf_coordinates", color = "black", size = 2.5) +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the people that test positive die from COVID-19.

```{r case/death}
ggplot(data=us, aes(x = cases/1000, y = deaths/1000)) + 
  geom_smooth(color = "black", size = 1.5, method = "loess", formula = y ~ x) +
  geom_jitter(color = "blue", size = .75) +
  theme_bw()
```

This plot shows how the cases and deaths relate and compare. I divided both cases and deaths by 1000 so its per thousand cases and per thousand deaths. 

### New York

```{r newyorkc}
ggplot(data = prop_county_ny) +
  geom_sf(aes(geometry = geometry), fill = NA,  color = "Black") +
  guides(color = FALSE) + 
  geom_text(aes(label = (as.numeric(death_case)*100), geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 2) +
  theme_bw() + 
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
```

This plot shows the percent of the people that test positive die from COVID-19 in the counties of New York.

## Data Tables

### States
```{r tibblea}
prop_state %>% 
  select(state, cases, deaths, pop, case_death, case_pop, death_pop, death_pop, death_case) %>% 
datatable(caption = 'This Tibble shows COVID-19 data in the states in the USA for yesterday',
          class = 'cell-border stripe', 
          rownames = FALSE, 
          filter = 'top', 
          extensions = 'Buttons',
          options = list(buttons = c('copy', 'csv', 'pdf'), 
                         pageLength = 5, 
                         autoWidth = TRUE, 
                         dom = 'Bfrtip'))
```

### Past Two Weeks

```{r tibbleb}
us %>% filter_biweek() %>% 
datatable(caption = 'This Tibble shows COVID-19 data in the USA for the last 2 weeks',
          class = 'cell-border stripe', 
          rownames = FALSE, 
          filter = 'top', 
          extensions = 'Buttons',
          options = list(buttons = c('copy', 'csv', 'pdf'), 
                         pageLength = 5, 
                         autoWidth = TRUE, 
                         dom = 'Bfrtip'))
```

<br /><br />

```{r update}
update()
```