---
title: "Our World in Data"
---

```{r setup, include=FALSE}
pacman::p_load(pacman, package)

libraries()

wdset("D:/COVID-project/")

owid <- my_read("./read-data/OWID/owid_data.csv") 
ecdc <- my_read("./read-data/OWID/owid_ecdc.csv") 
```

```{r wrangling}
continent <- owid %>% 
  group_by(continent, date) %>% 
  my_owid_wrangle()

yesterday_continent <- continent %>% 
  filter(date >= Sys.Week()) 

world <- owid %>% 
  group_by(date) %>% 
  my_owid_wrangle()

yesterday_world <- world %>% 
  filter(date >= Sys.Week()) 
```

## Relationship with Continents

### Cases

```{r casesa}
ggplot(data = continent, aes(x = date, y = total_cases/1000, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows how total cases have increased over time colored by continent. I divided total cases by one thousand so it's per one thousand cases. 

```{r casesb}
ggplot(data = continent, aes(x = continent, y = total_cases/1000000, fill = continent)) +
  geom_col() + theme_bw()
```

This plot shows total over all cases by continent. I divided total cases by one million so it's per one million cases. 

```{r casesc}
ggplot(data = continent, aes(x = date, y = new_cases/1000, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows new cases each day colored by continent. I divided new cases by one thousand so it's per one thousand cases.

### Deaths

```{r deathsa}
ggplot(data = continent, aes(x = date, y = total_deaths/1000, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows how total deaths have increased over time colored by continent. I divided total deaths by one thousand so it's per one thousand deaths. 

```{r deathsb}
ggplot(data = continent, aes(x = continent, y = total_deaths/1000, fill = continent)) +
  geom_col() + theme_bw()
```

This plot shows total over all deaths by continent. I divided total deaths by one thousand so it's per one thousand deaths. 

```{r deathsc}
ggplot(data = continent, aes(x = date, y = new_deaths/100, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows new deaths each day colored by continent. I divided new deaths by one hundred so it's per one hundred deaths.

### Tests

```{r testsa}
ggplot(data = continent, aes(x = date, y = total_tests/1000000, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows how total tests have increased over time colored by continent. I divided total tests by one million so it's per one million tests. 

```{r testsb}
ggplot(data = continent, aes(x = continent, y = total_tests/1000000, fill = continent)) +
  geom_col() + theme_bw()
```

This plot shows total over all tests by continent. I divided total tests by one million so it's per one million tests. 

```{r testsc}
ggplot(data = continent, aes(x = date, y = new_tests/1000, color = continent)) +
  geom_line() + theme_bw()
```

This plot shows new tests each day colored by continent. I divided new tests by one thousand so it's per one thousand tests.

## Relationship with the World

### Cases

```{r caseswa}
ggplot(data = world, aes(x = date, y = total_cases/1000)) +
  geom_line() + theme_bw()
```

This plot shows how total cases have increased over time for the whole world. I divided total cases by one thousand so it's per one thousand cases. 

```{r caseswb}
ggplot(data = world, aes(x = date, y = new_cases/1000)) +
  geom_line() + theme_bw()
```

This plot shows new cases each day for the whole world. I divided new cases by one thousand so it's per one thousand cases.

### Deaths

```{r deathswa}
ggplot(data = world, aes(x = date, y = total_deaths/1000)) +
  geom_line() + theme_bw()
```

This plot shows how total deaths have increased over time for the whole world. I divided total deaths by one thousand so it's per one thousand deaths. 

```{r deathswb}
ggplot(data = world, aes(x = date, y = new_deaths/100)) +
  geom_line() + theme_bw()
```

This plot shows new deaths each day for the whole world. I divided total deaths by one hundred so it's per one hundred deaths.

### Tests

```{r testswa}
ggplot(data = world, aes(x = date, y = total_tests/1000000)) +
  geom_line() + theme_bw()
```

This plot shows how total tests have increased over time for the whole world. I divided total tests by one million so it's per one million cases. 

```{r testswb}
ggplot(data = world, aes(x = date, y = new_tests/1000)) +
  geom_line() + theme_bw()
```

This plot shows new tests each day for the whole world. I divided total tests by one thousand so it's per one thousand tests.

<br /><br />

## Data Tables

### Continents

```{r tibblea}
yesterday_continent %>% 
datatable(caption = 'This Tibble shows COVID-19 data by continents for the last 2 weeks',
          class = 'cell-border stripe', 
          rownames = FALSE, 
          filter = 'top', 
          extensions = 'Buttons',
          options = list(buttons = c('copy', 'csv', 'pdf'), 
                         pageLength = 5, 
                         autoWidth = TRUE, 
                         dom = 'Bfrtip'))
```

### World

```{r tibbleb}
yesterday_world %>% 
datatable(caption = 'This Tibble shows COVID-19 data for the last 2 weeks in the world',
          class = 'cell-border stripe', 
          rownames = FALSE, 
          filter = 'top', 
          extensions = 'Buttons',
          options = list(buttons = c('copy', 'csv', 'pdf'), 
                         pageLength = 5, 
                         autoWidth = TRUE, 
                         dom = 'Bfrtip'))
```

<br /><br />

```{r update}
update
```