---
title: "Progress Report"
subtitle: "Week 3"
---

```{r load_libraries, include=FALSE}
package::libraries()

dt <- read_csv("./time_logs/pr03.csv")
```

## Work Schedule --- Updated

Below I have divided the tasks in three-week interval which a line progress reports. 

### Progress Report 2 (Due 5/30)

* Load in the data sets
  + Look through the data set
* Data_Push R package  
* Start wrangling
* Make basic visualizations to see the data
* Develop YML file 

### Progress Report 3 (Due 6/20)

* Wrangle and Visualize data
  + Non-interactive plots
  + Interactive plots

### Progress Report 4 (Due 7/11)

* Last steps website
  + Finalize visualizations 
  + Finalize YML
  + Finalize tibbles
* Start report

### Final Delivery (7/22)

* Finish data presentation website
* Peer reviewed 
  + Make edits/modifications
* Complete report

### Whole Semester (4/20 - 7/22)

Throughout the semester I will spend time gaining knowledge on the follow as it relates to my project:  

&emsp; (Notes and Resources will be posted in the organizations github repos)

* Ethics in Data
* Consulting in Statistics and Data Science
* Analysis - Poor, [Good, Better, Best](https://www.churchofjesuschrist.org/study/general-conference/2007/10/good-better-best?lang=eng)
* R packages
* Making Inferences with Data
* Learn more about the 
  + [ggplot](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.0)
  + [sf plots](https://r-spatial.github.io/sf/articles/sf1.html)
  + [leaflet](https://leafletjs.com/)
  + [dygraphs](https://rstudio.github.io/dygraphs/)
  + other plot tools
* and more

## Accomplishments 

### Progress Report 1 (Due 5/9)

* Start Project
  + Proposal
* Do some data research
* Find data sets to work with
  + Such as [JHUs github](https://github.com/CSSEGISandData/COVID-19)
  + Or [CDC](https://www.cdc.gov/flu/weekly/weeklyarchives2019-2020/data/whoAllregt_phl15.html) flu data
* Create github organization
* Host the website in main repo

### Hour Break Down

```{r display}
kable(dt)
```

## Observations 

In my data research for this COVID-19 Data Project I have come across many useful data sets. However, because Git hub restricts files to be no larger than 100MB, another problem arises, most of the 100MB will be taken up by reading in the different data sets. My proposed solution is to create my own data library. Be assured this problem and acquired solution wouldn't effect my work schedule.
