---
title: "Progress Report"
subtitle: "Week 9"
---

```{r libraries, include=FALSE}
package::libraries()

dt <- read_csv("./time_logs/pr09.csv")
```


## Work Schedule --- Updated

Below I have divided the tasks in three-week interval which a line progress reports. 

### Progress Report 4 (Due 7/11)

* Last steps website
  + Finalize visualizations 
  + Finalize YML
  + Finalize tibbles
* Start report

### Final Delivery (7/22)

* Finish data presentation website
* Peer reviewed 
  + Make edits/modifications
* Complete report

### Whole Semester (4/20 - 7/22)

Throughout the semester I will spend time gaining knowledge on the follow as it relates to my project:  

&emsp; (Notes and Resources will be posted in the organizations github repos)

* Ethics in Data
* Consulting in Statistics and Data Science
* Analysis - Poor, [Good, Better, Best](https://www.churchofjesuschrist.org/study/general-conference/2007/10/good-better-best?lang=eng)
* R packages
* Making Inferences with Data
* Learn more about the 
  + [ggplot](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.0)
  + [sf plots](https://r-spatial.github.io/sf/articles/sf1.html)
  + [leaflet](https://leafletjs.com/)
  + [dygraphs](https://rstudio.github.io/dygraphs/)
  + other plot tools
* and more

## Accomplishments 

### Progress Report 1 (Due 5/9)

* Start Project
  + Proposal
* Do some data research
* Find data sets to work with
  + Such as [JHUs github](https://github.com/CSSEGISandData/COVID-19)
  + Or [CDC](https://www.cdc.gov/flu/weekly/weeklyarchives2019-2020/data/whoAllregt_phl15.html) flu data
* Create github organization
* Host the website in main repo

### Progress Report 2 (Due 5/30)

* Load in the data sets
  + Look through the data set
* Data_Push R package (Data host repo)  
* Start wrangling
* Make basic visualizations to see the data
* Develop YML file 

### Progress Report 3 (Due 6/20)

* Wrangle and Visualize data
  + Non-interactive plots
  + Interactive plots

### Hour Break Down

```{r display, echo=FALSE}
kable(dt)
```

## Observations 

I was able to complete the framework of the YAML file. I was able to complete the visualizations for NYTimes and got started with both JHU and OWID. I did a mini project called tidy tuesday about sleep patterns. I was able to learn more about Tableau and get Data. I learned about the R package installr. I also learned more about data-driven learning and research. I believe still that I am on schedule to complete the project by the end of the week. 
